### 2013-09-18 intro.tex
state of the art, slight copy-editing....

\chapter{Introduction}
This thesis investigates the use of structured learning algorithms to the task
of semantic image segmentation. Both topics have recieve much attention in the
computer vision and machine learning communities lately. Unfortunately,
learning structured prediction in computer vision applications is still little
understood.  We hoe this thesis will further the understanding of structured
learning methods, give practical guidance and provide implementations to
further the use of structured prediction methods in computer vision.

The task of semantic segmentation is defined as predicting a semantic class label
for each pixel in a natural image. Predicting semantic labels on such a fine scale
constitutes an important step towards general scene understanding.
%
Applications of scene understanding include automatic image interpretation for retrieval,
 autonomous driving and mobile robotics.
%
related work?? here again?
Current systems for semantic segmentation emply a wealth of different paradigms, most notably:
The ranking of figure ground hypotheses
conditional random field / superpixel? kraehenbuehl?
detector based
sparse coding

We will focus on the use of conditional random fields, which have shown very
promising.  Using the paradigm of structural support vector machines, it is
possible to learn conditional random fields to directly minimize a loss of
interest. In particular, conditional random fields allow to combine different
cues, possibly produced by other approaches, in a principled manner.
%
We use four different datasets to evaluate our approach: Graz02, Pascal VOC
2010, MSRC-21 and NYU V2.
%
Why semantic segmentation? Difference object / stuff centric? -> difference to Alex
Structured approaches
%

\section{List of Contributions}
Contributions:
\begin{itemize}
\item Introduces general clustering algorithm that improves pipeline in serveral places.
\item Shows that weakly supervised object class segmentation (not semantic segmentation) is possible.
\item Provides general efficient framework for structured prediction.
\item Analyze max-margin learning algorithms in different regimes.
\item shows exact learning is possible in loopy graphs and improves results (sometimes?).
\item shows importance of edge features.
\end{itemize}

\section{Thesis Outline}
We start by introducing a novel clustering algorithm in Chapter~\ref{ch:itm},
which is used as a building block in several places for our semantic
segmentation algorithms.  We will then introducing a mostly unstructured
approach to semi-supervised learning for semantic segmentation in
Chaper~\ref{ch:semi_supervised}, investigating the problem of annotating
training data for semantic segmentation.  The central topic of this thesis is
introduced in Chapter~\ref{ch:structured_pystruct}, which also introduces our
software library for implementing structured learning and prediction
algorithms.
Chapter~\ref{ch:comparision} gives a quantitative comparison of the most widely
used structured prediction algorithms, focusing on which algorithm is best to
use in a given situation. In particular, we investigate learning behavior for
semantic segmentation on several datasets.

The problem of learning with approximate inference is investigated in
Chapter~\ref{ch:exact_learning}.  We develop a strategy for efficient caching
and a combination of inference algorithms that allows us to learn SSVMs for
semantic image segmentation exactly, even though the involved factor graphs
contain many loops.

Finally, Chapter~\ref{ch:nyu} applies the methods developed in
Chapter~\ref{ch:structurd_pystruct} and Chapter~\ref{ch:exact_learning} to the
problem of semantic annotation with structure classes in RGB-D data. We
demonstrate that we are able to learn meaningful spacial relations, and improve
upon the state of the art in the NYU V2 datasets that we consider.

>>>>

\chapter{Introduction}
The task of semantic segmentation is defined as predicting a semantic class label
for each pixel in a natural image. Predicting semantic labels on such a fine scale
constitutes an important step towards general scene understanding.
%
Applications of scene understanding include automatic image interpretation for retrieval,
autonomous driving and mobile robotics.
%
We will in the following distinguish between the task of semantic segmentation,
which is usually distinguishing unstructured ``stuff'' classes such as road and
grass, and object class segmentation, which denotes the segmentation of very
structured classes, such as cars, planes and people.
We consider four different datasets in this thesis: the object class segmentation
datasets Graz02 and Pascal VOC 2010, and the semantic segmentation datasets MSRC-21,
and NYU V2.
%
Both tasks have the same ultimate goal of parsing, and therefore understanding, images
in terms of semantic classes, but usually employ  different mechanisms to represent the input.
%
%The main part of this thesis investigates the tasks of semantic segmentation and object class segmentation in natural images.
The main part of this thesis investigates the use of structured learning
algorithms to the task of semantic image segmentation. Both topics have receive
much attention in the computer vision and machine learning communities lately.
Unfortunately, learning structured prediction in computer vision applications
is still little understood.  We hoe this thesis will further the understanding
of structured learning methods, give practical guidance and provide
implementations to further the use of structured prediction methods in computer
vision.

%related work?? here again?
%Current systems for semantic segmentation emply a wealth of different paradigms, most notably:
%The ranking of figure ground hypotheses
%conditional random field / superpixel? kraehenbuehl?
%detector based
%sparse coding

We will focus on the use of conditional random fields, which have shown very
promising.  Using the paradigm of structural support vector machines, it is
possible to learn conditional random fields to directly minimize a loss of
interest. In particular, conditional random fields allow to combine different
cues, possibly produced by other approaches, in a principled manner.
%
Why semantic segmentation? Difference object / stuff centric? -> difference to Alex
Structured approaches
%

\section{List of Contributions}
\begin{itemize}
\item Introduction of general clustering algorithm that improves upon widely used
    approaches from the literature.
\item Demonstration of an algorithm for  weakly supervised object class segmentation.
\item Providing a general efficient implementation for structured prediction.
\item Analysis of max-margin learning algorithms with exact or approximate inference in different applications.
\item Demonstrating that exact learning for semantic segmentation and object class segmentation is possible in loopy graphs.
%\item Shows importance of edge features.
\end{itemize}

\section{Thesis Outline}
We start by introducing a novel clustering algorithm in Chapter~\ref{ch:itm},
which we plan to use as an additional building block for semantic segmentation
algorithms in future work. Clustering is important both for the initial
segmentation into superpixels, and in the creation of visual words.

We will then introducing a mostly unstructured
approach to semi-supervised learning for object class segmentation in
Chaper~\ref{ch:semi_supervised}, investigating the problem of annotating
training data for semantic segmentation. The central topic of this thesis is
introduced in Chapter~\ref{ch:structured_pystruct}, which also introduces our
software library for implementing structured learning and prediction
algorithms.
Chapter~\ref{ch:comparision} gives a quantitative comparison of the most widely
used structured prediction algorithms, focusing on which algorithm is best to
use in a given situation. In particular, we investigate learning behavior for
semantic segmentation on several datasets.

The problem of learning with approximate inference is investigated in
Chapter~\ref{ch:exact_learning}.  We develop a strategy for efficient caching
and a combination of inference algorithms that allows us to learn SSVMs for
semantic image segmentation exactly, even though the involved factor graphs
contain many loops. We demonstrate our algorithm on the Pascal VOC 2010 and MSRC-21 datasets,
and provide state-of-the-art results on the MSRC-21 dataset.

Finally, Chapter~\ref{ch:nyu} applies the methods developed in
Chapter~\ref{ch:structurd_pystruct} and Chapter~\ref{ch:exact_learning} to the
problem of semantic annotation with structure classes in RGB-D data. We
demonstrate that we are able to learn meaningful spacial relations, and improve
upon the state-of-the-art in the NYU V2 datasets that we consider.

### 2013-09-18 main.tex
state of the art, slight copy-editing....

\begin{center}
\Large\sffamily
\vspace*{1cm}
\textbf{{\huge Learning Structured Prediction For Semantic Image Segmentation}}\\[5mm]
Dissertation\\%[3mm]
zur\\%[3mm]
Erlangung des Doktorgrades (Dr.\ rer.\ nat.)\\%[3mm]
der\\%[3mm]
Mathematisch-Naturwissenschaftlichen Fakult\"at\\%[3mm]
der\\%[3mm]
Rheinischen Friedrich-Wilhelm Universit\"at Bonn\\%[3mm]
vorgelegt von\\%[5mm]
{\LARGE Andreas Christian M\"uller}\\%[5mm]
aus\\%[3mm]
Offenbach am Main\\%[10mm]
Bonn, September 2013
\end{center}
\end{titlepage}
%
%\thispagestyle{empty}
\clearpage
\pagenumbering{Roman}
\setcounter{page}{2}
\vspace*{5cm}
\Large\sffamily
\begin{center}
Angefertigt mit Genehmigung\\
der Mathematisch-Naturwissenschaftlichen Fakult\"at\\
der Rheinischen Friedrich-Wilhelms-Universit\"at Bonn\\
\end{center}
1. Gutachter Prof. Dr. Sven Behnke\\
2. Gutachter Prof. Dr. Reinhard Klein\\
Tag der Promotion:\\
Erscheinungsjahr: 2013\\
\end{spacing}
\tableofcontents
%\listoffigures
%\listoftables
\chapter*{Zusammenfassung}
Automatische Segementierung und Erkennung von semantischen Klassen in
nat\"urlichen Bildern ist ein wichtiges offenes Problem des machinellen Sehens.
Diese Arbeit untersucht drei m\"oglichen Ansi\"atzen der Erkennung:
ohne \"Uberwachung, mit \"Uberwachung auf Ebene von Bildern und mit \"Uberwachung auf Ebene
von Pixeln. Diese Arbeit setzt sich aus drei Teilen zusammen.

Im ersten Teil der Arbeit schlagen wir einen Clustering-Algorithmus vor,
der eine neuartige, informationstheoretische Zielfunktion optimiert. Wir zeigen, dass
der vorgestellte Algorithmus \"ublichen Standardverfahren aus der Literatur gegen\"uber
klare Vorteile auf vielen verschiedenen Datens\"atzen hat.

@@

Der dritte und letzte Teil er Arbeit untersucht einen der am weitest
verbreiteten Ans\"atze zu semantischen Segmentierung und
Objektklassensegmentierung, n\"amlich Conditional Random Fields verbunden mit Verfahren
der strukturierten Vorhersage.
%Die Vorhersage f\"ur Modelle zur Bild-Markierung
%mittels Conditional Random Fields ist f\"ur gew\"ohnlich nur approiximativ
%moeglich, bedingt d\"urch das Vorhandensein von Kreisen in den zugrunde
%liegenden Nachbarschaftsgraphen.  Aus diesem Grund, und trotz der Beliebtheit
%von Conditional Random Field Verfahren, ist das strukturierte Lernen f\"ur
%solche Anwendungen schlecht verstanden.
%
Wir untersuchen verschiedene Lernalgorithmen des struktuierten Lernens, insbesondere
im Zusammenhang mit approximativer Inferenz. Wir zeigen, dass es m\"oglich ist trotz des Vorhandenseins von
Kreisen in den Nachbarschaftsgraphen exakte strukturierte Modelle zur Bildsegmentierung zu lernen.
Wir stellen ausserdem eine Software-Bibliothek vor, die es erlaubt einen
weitreichenden Vergleich der besten Lernverfahren f\"ur strukturiertes Lernen
durchzuf\"uhren.
Unsere Studie erlaubt uns eine Charakterisation der betrachteten Algorithmen
in einer Reihe von Anwendungen, insbesondere der semantischen Segmentierung und
Objektklassensegmentierung.

@@

The first part introduces a clustering algorithm, that optimizes a novel information-theoretic
objective function. We show that the proposed algorithm has clear advantages over standard algorithm
from the Literature on a wide array of datasets.

@@

The third and last part of this work investigates one of the most popular
approaches to the task of object class segmentation and semantic segmentation,
based on conditional random fields and structured prediction.
%
%For image labeling tasks, prediction using conditional random fields is usually
%only approximate, because of the presence of loops of the neighborhood graph in
%the the image. Consequently, learning in this setting is still little
%understood, despite the popularity of these methods.
%
We investigate several learning algorithms, in particular in combination with
approximate inference procedures. We show how structured models for image
segmentation can be learned exactly in practical settings, even in the presence
of many loops in the underlying neighborhood graphs.  We introduce a software
library that allows us to perform extensive empirical comparisons of
state-of-the-art structured learning approaches. This allows us to characterize
their practical properties in a range of applications, in particular for
semantic segmentation and object class segmentation.

@@

I have been lucky to be a part of the scikit-learn community, whose developers
have been a constant source of encouragement and inspiration to me. My
understanding of machine learning as well as software engineering has been
greatly affected by the continuing exchange with Olivier Grisel, Lars Buitinck,
Gael Varoquax, Vlad Niculae, Mathieu Blondel, Gilles Louppe, Arnaud Joly and
others.

@@

Last but not least I am greateful to Anna M\"uller for her support,
encouragement and company. Thank you for giving me the strength needed for this
endeavour.
\cleardoublepage
\pagenumbering{arabic}
\include{intro}

@@

%\include{itm/paper}

%\include{partial_supervision/paper}

@@

\chapter{Conclusion and Future Direction}
Current research in computer vision progresses at an incredible speed, with new
methods and applications emerging constantly.  Still, there are large
differences in the paradigms employed by different researchers.  While it is
generally acknowledged that context is one of the most important cues in
recognizing objects and interpreting scenes, there is little consensus on how
contextual information should be modelled, or even learned.
The structured prediction paradigm allows a principled approach to
incorporating predictions within an image, and to include context information.
We investigated the potential of structural methods for the application of
semantic image segmentation and object recognition, applications to which
structured methods have been applied successfully before.  We focus on
\emph{learning} of structural models and the interaction of inference and
learning in the neighborhood models typically employed for semantic
segmentation.
The thesis presents a software implementation of a variety of popular learning
algorithms for structural support vector machines, together with a thorough
evaluation of their properties in the presence of approximate inference.
We demonstrate that exact learning is possible, even in the presence of loops in the
underlying factor graph. We also demonstrate the power of conditioning interactions
on data, learning spacial interactions in a RGB-D setting.

>>>>

\begin{center}
\Large\sffamily
\vspace*{1cm}
\textbf{{\huge Learning Structured Prediction For Semantic Image Segmentation}}\\[5mm]
Dissertation\\%[3mm]
zur\\%[3mm]
Erlangung des Doktorgrades (Dr.\ rer.\ nat.)\\%[3mm]
der\\%[3mm]
Mathematisch-Naturwissenschaftlichen Fakult\"at\\%[3mm]
der\\%[3mm]
Rheinischen Friedrich-Wilhelm Universit\"at Bonn\\%[3mm]
vorgelegt von\\%[5mm]
{\LARGE Andreas Christian M\"uller}\\%[5mm]
aus\\%[3mm]
Offenbach am Main\\%[10mm]
Bonn, September 2013
\end{center}
\end{titlepage}
%
%\thispagestyle{empty}
\clearpage
\pagenumbering{Roman}
\setcounter{page}{2}
\vspace*{5cm}
\Large\sffamily
\begin{center}
Angefertigt mit Genehmigung\\
der Mathematisch-Naturwissenschaftlichen Fakult\"at\\
der Rheinischen Friedrich-Wilhelms-Universit\"at Bonn\\
\end{center}
1. Gutachter Prof. Dr. Sven Behnke\\
2. Gutachter Prof. Dr. Reinhard Klein\\
Tag der Promotion:\\
Erscheinungsjahr: 2013\\
\end{spacing}
\tableofcontents
%\listoffigures
%\listoftables
\chapter*{Zusammenfassung}
\enlargethispage{10mm}
Automatische Segmentierung und Erkennung von semantischen Klassen in
nat\"urlichen Bildern ist ein wichtiges offenes Problem des machinellen Sehens.
Diese Arbeit untersucht drei m\"oglichen Ansi\"atzen der Erkennung:
ohne \"Uberwachung, mit \"Uberwachung auf Ebene von Bildern und mit \"Uberwachung auf Ebene
von Pixeln. Diese Arbeit setzt sich aus drei Teilen zusammen.

Im ersten Teil der Arbeit schlagen wir einen Clustering-Algorithmus vor,
der eine neuartige, informationstheoretische Zielfunktion optimiert. Wir zeigen, dass
der vorgestellte Algorithmus \"ublichen Standardverfahren aus der Literatur gegen\"uber
klare Vorteile auf vielen verschiedenen Datens\"atzen hat. Clustering ist ein wichtiger
Baustein in vielen Applikationen des Machinellen Sehens, insbesondere in der
automatischen Segmentierung.

@@

Der dritte und letzte Teil er Arbeit untersucht einen der am weitest
verbreiteten Ans\"atze zu semantischen Segmentierung und
Objektklassensegmentierung, n\"amlich Conditional Random Fields verbunden mit Verfahren
der strukturierten Vorhersage.
%Die Vorhersage f\"ur Modelle zur Bild-Markierung
%mittels Conditional Random Fields ist f\"ur gew\"ohnlich nur approiximativ
%moeglich, bedingt d\"urch das Vorhandensein von Kreisen in den zugrunde
%liegenden Nachbarschaftsgraphen.  Aus diesem Grund, und trotz der Beliebtheit
%von Conditional Random Field Verfahren, ist das strukturierte Lernen f\"ur
%solche Anwendungen schlecht verstanden.
%
Wir untersuchen verschiedene Lernalgorithmen des struktuierten Lernens, insbesondere
im Zusammenhang mit approximativer Inferenz. Wir zeigen, dass es m\"oglich ist trotz des Vorhandenseins von
Kreisen in den Nachbarschaftsgraphen exakte strukturierte Modelle zur Bildsegmentierung zu lernen.
Mit den vorgestellten Methoden bringen den Stand der Kunst auf
auf zwei komplexen Datens\"atzen zur semantischen Segmentierung voran, dem MSRC-21 Datensatz
von RGB-Bildern und dem NYU V2 Datensatz von RGB-D Bildern von Innenraum-Szenen.

Wir stellen ausserdem eine Software-Bibliothek vor, die es erlaubt einen
weitreichenden Vergleich der besten Lernverfahren f\"ur strukturiertes Lernen
durchzuf\"uhren.
Unsere Studie erlaubt uns eine Charakterisation der betrachteten Algorithmen
in einer Reihe von Anwendungen, insbesondere der semantischen Segmentierung und
Objektklassensegmentierung.

@@

The first part introduces a clustering algorithm, that optimizes a novel information-theoretic
objective function. We show that the proposed algorithm has clear advantages over standard algorithm
from the Literature on a wide array of datasets. Clustering algorithms are an important building block
for higher level computer vision applications, in particular for semantic segmentation.

@@

The third and last part of this work investigates one of the most popular
approaches to the task of object class segmentation and semantic segmentation,
based on conditional random fields and structured prediction.
%
%For image labeling tasks, prediction using conditional random fields is usually
%only approximate, because of the presence of loops of the neighborhood graph in
%the the image. Consequently, learning in this setting is still little
%understood, despite the popularity of these methods.
%
We investigate several learning algorithms, in particular in combination with
approximate inference procedures. We show how structured models for image
segmentation can be learned exactly in practical settings, even in the presence
of many loops in the underlying neighborhood graphs.
The introduced methods provide results advancing the state-of-the art on two complex benchmark datasets
for semantic segmention, the MSRC-21 Dataset of RGB images and the NYU V2 Dataset or RGB-D
images of indoor scenes.

We also introduce a software library that allows us to perform extensive empirical
comparisons of state-of-the-art structured learning approaches. This allows us
to characterize their practical properties in a range of applications, in
particular for semantic segmentation and object class segmentation.

@@

I have been lucky to be a part of the scikit-learn community, whose developers
have been a constant source of encouragement and inspiration to me. My
understanding of machine learning as well as software engineering has been
greatly affected by the continuing exchange with Olivier Grisel, Lars Buitinck,
Ga\"el Varoquax, Vlad Niculae, Mathieu Blondel, Gilles Louppe, Arnaud Joly and
others.

@@

Last but not least I am grateful to Anna M\"uller for her support,
encouragement and company. Thank you for giving me the strength needed for this
endeavour.
\cleardoublepage
\pagenumbering{arabic}
\include{intro}

@@

\include{itm/paper}

\include{partial_supervision/paper}

@@

\chapter{Conclusion and Future Direction}
Current research in computer vision progresses at an incredible speed, with new
methods and applications emerging constantly.  Still, there are large
differences in the paradigms employed by different researchers.  While it is
generally acknowledged that context is one of the most important cues in
recognizing objects and interpreting scenes, there is little consensus on how
contextual information should be modelled, or even learned.
The structured prediction paradigm allows a principled approach to
incorporating predictions within an image, and to include context information.
We investigated the potential of structural methods for the application of
semantic image segmentation and object recognition, applications to which
structured methods have been applied successfully before.  We focus on
\emph{learning} of structural models and the interaction of inference and
learning in the neighborhood models typically employed for semantic
segmentation.
The thesis presents a software implementation of a variety of popular learning
algorithms for structural support vector machines, together with a thorough
evaluation of their properties in the presence of approximate inference.
We demonstrate that exact learning is possible, even in the presence of loops in the
underlying factor graph. We also demonstrate the power of conditioning interactions
on data, learning spacial interactions in a RGB-D setting.
%TODO STATE OF THE F*ING ART!
